---
title: "R_figure_generator_Supplements"
author: "Michiel"
date: '2023-06-29'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# GENERAL SETUP

```{r}
rm(list = ls())
cat('\014')
pc_name <- file.path('C:/users', Sys.info()[7])
setwd(file.path(pc_name, 'Dropbox (Verstrepen.lab)', 'BeerAromas', 'submission data final', 'Supplements'))
```

# Bar plot showing the number of beers for each style

```{r}
library(tidyverse)
```

### Loading in data

```{r}
data_RB <- read.csv('Supplemental File S1.csv') %>% 
  mutate(tasting_category_fine = factor(tasting_category_fine, levels = rev(sort(unique(tasting_category_fine)))))

data_RB <- data_RB %>% 
  select(beer, tasting_category_fine) %>% 
  group_by(tasting_category_fine) %>% 
  summarise(total = n()) %>% 
  mutate(tasting_category_fine = gsub('Low/No alcohol', 'Non/low-alcohol', tasting_category_fine), 
         tasting_category_fine = factor(tasting_category_fine, levels = rev(c('Pils/Lager',
                                                                              'Wheat', 
                                                                              'Saison', 
                                                                              'Blond', 
                                                                              'Tripel',
                                                                              'Strong ale',
                                                                              'Amber', 
                                                                              'Brown', 
                                                                              'Dubbel',
                                                                              'Stout/Porter', 
                                                                              'Scotch', 
                                                                              'Christmas', 
                                                                              'Hoppy', 
                                                                              'Flanders old brown', 
                                                                              'West Flanders ale',
                                                                              'Faro', 
                                                                              'Lambic', 
                                                                              'Kriek', 
                                                                              'Fruitbeer', 
                                                                              'Brett/cofermented',
                                                                              'Brut', 
                                                                              'Non/low-alcohol'
  )))) %>% 
  mutate(label = paste(total, ' ', sep = ''))
```

### Plotting

Plot settings

```{r}
{
  fill = '#Faca13'
  y_ax <- 'Number of beers'
  y_angle <- 90
  x_ax <- 'Beer style'
  textsize = 17
  pointsize = 1.5
  legend_size = 0.85
  textcolor = '#000000'
}
```

Bar plot

```{r}
style_plot_no_whitespace <- ggplot(data_RB, aes(x = tasting_category_fine, y = total)) + 
  geom_bar(fill = fill, stat = 'identity') + 
  geom_text(aes(label = label), hjust = 1, fontface= 'bold', color = 'white', size = textsize/3) + 
  theme_bw() +
  labs(x = x_ax,
       y = y_ax,
       color = "Beer style") +
  theme(axis.title.y = element_text(angle = y_angle, size = round(18*textsize/15), color = textcolor, face = 'bold'),
        axis.title.x = element_text(size = round(18*textsize/15), color = textcolor, face = 'bold'),
        axis.text.x = element_text(size = textsize, color = textcolor),
        axis.text.y = element_text(size = textsize, color = textcolor),
        axis.line = element_line(color = textcolor),
        axis.ticks = element_line(color = textcolor),
        plot.title = element_text(size = round(18*textsize/15)),
        legend.title = element_text(size = round(16*textsize/15), color = textcolor),
        legend.text = element_text(size = round(16*textsize/15), color = textcolor),
        legend.key.size = unit(legend_size, 'cm'),
        panel.grid = element_blank(),
        strip.text = element_text(size = round(16*textsize/15), face = 'bold')) +
  ggtitle("") + 
  scale_y_continuous(expand = expansion(add = c(0, 1))) +
  coord_flip()

ggsave(plot = style_plot_no_whitespace,
       filename = 'R_Beer_per_style_no_whitespace.png',
       units = 'mm',
       height = 200,
       width = 250,
       dpi = 300)
```

# Beer appreciation vs beer price

Note: Because the RateBeer scores are property of RateBeer (ZX Ventures USA), only the data for the trained panel dataset is provided for this Figure.

```{r}
library(tidyverse)
library(tidyquant)
library(ggpmisc)
library(svglite)
```

### Loading data

```{r}
prices <- read.csv('Beer Prices.csv')
# data_RB <- read.csv('Submission data final/chem_dataset_complete_log_ratebeer.csv')
data_EP <-  read.csv('Supplemental File S4.csv')
```

### Data wrangling

```{r}

# data_RB <- data_RB %>% 
#   select(beer, beer_id, overall_20) %>% 
#   rename(RateBeer = overall_20)

data_EP <- data_EP %>% 
  select(beer, beer_id, overall) %>% 
  rename(`Trained panel` = overall)


colruyt_delhaize <- prices %>% 
  select(beer, beer_id, tasting_category, Price_Colruyt_Belgium, Price_Delhaize_Belgium) %>% 
  rename(Colruyt = Price_Colruyt_Belgium,
         Delhaize = Price_Delhaize_Belgium) %>% 
  merge(data_EP, all = TRUE) %>% 
  # merge(data_RB, all = TRUE) %>% 
  pivot_longer(cols = c(Colruyt,
                        Delhaize),
               names_to = 'store', values_to = 'price') %>% 
  # pivot_longer(cols = c(RateBeer, `Trained panel`), names_to = 'dataset', values_to = 'appreciation') %>% 
  mutate(store = case_when(store == 'Colruyt' ~ 'Colruyt price ($)',
                           store == 'Delhaize' ~ 'Delhaize price ($)',
                           TRUE ~ store)) %>% 
  filter(!is.na(price)) %>% 
  mutate(dataset = "Trained panel") %>% 
  rename(appreciation = "Trained panel")

correlations <- colruyt_delhaize %>% 
  group_by(store, dataset) %>% 
  summarise(correlation = cor(price, appreciation, method = 'spearman'),
            price = 8,
            appreciation = -2,
            n = n())
```

### Plotting

```{r}
{
  y_ax <- 'Appreciation'
  y_angle <- 0
  x_ax <- 'Price'
  viridis_code = 'C'
  textsize = 17
  pointsize = 1.5
  legend_size = 0.85
  textcolor = '#192945'
}

CD <- ggplot(colruyt_delhaize, aes(x = appreciation, y = price)) +
  facet_grid(cols = vars(dataset),
             rows = vars(store),
             #scales = 'free',
             switch = 'both') + 
  geom_text(data = correlations,
            aes(label = paste0("ρ = ", round(correlation, 2))),
            size = 5,
            color = textcolor) + 
  geom_point(size = 2, alpha = 0.7, color = textcolor) +
  theme_bw() +
  labs(x = 'Appreciation score',
       y = '') +
  theme(axis.title.y = element_text(size = round(18*textsize/15), color = textcolor),
        axis.title.x = element_text(size = round(18*textsize/15), color = textcolor),
        axis.text.y = element_text(size = round(16*textsize/15), color = textcolor),
        axis.text.x = element_text(size = round(16*textsize/15), color = textcolor),
        axis.line = element_blank(),
        axis.ticks = element_line(color = textcolor),
        plot.title = element_text(size = round(18*textsize/15)),
        legend.title = element_text(size = round(16*textsize/15), color = textcolor),
        legend.text = element_text(size = round(16*textsize/15), color = textcolor),
        legend.position = 'none',
        legend.key.size = unit(legend_size, 'cm'),
        panel.grid = element_blank(),
        strip.text = element_text(size = round(16*textsize/15),
                                  color = textcolor),
        strip.background = element_blank(),
        strip.placement = 'outside') +
  ggtitle("")


ggsave(filename = 'Appreciation_vs_price.svg',
       plot = CD,
       height = 165,
       width = 150,
       units = 'mm',
       dpi = 300)
```

# Rugplots

Note: Because the RateBeer scores are property of RateBeer (ZX Ventures USA), only the data for the trained panel dataset is provided for this Figure.

Libraries

```{r}
library(tidyverse)
library(RColorBrewer)
library(gridExtra)
library(ggpubr)
```

Colors

```{r}
color_yeast <- "#E86A3B"
color_malt <- "#6BA8C9"
color_hop <- "#ABD17D"
color_wildflora <- "#FACA13"
color_other <- "#5A5F62"
palette <- c(Malt = color_malt, 
             Hops = color_hop, 
             Yeast = color_yeast, 
             Spices = color_other, 
             Others = color_other,
             "Wild flora" = color_wildflora)
color_beer <- "#E69F00"
color_darkbeer <- "#68480B"

print <- TRUE
```

### Gathering datasets

```{r}
chem_calibrated = read.csv('Calibrated chemical dataset.csv')

sensory_data = read.csv("Supplemental File S4.csv")

mapping_columnnames <- read.csv("Sensory names.csv")

thresholds <- read.csv("Sensory thresholds.csv")
```

### Chemical Rugplots

We create a nested dataframe that contains all the data we need for plotting.

1.  Setting the compounds that we want to plot
2.  Adding all the other required info
3.  Creating the 'clean_name' parameter

```{r}
rugplot_data = rbind(tibble(plotgroup = 'group 1',
                            field_name_R = c("ethyl_acetate", "ethyl_decanoate", "ethyl_hexanoate", 
                                             "isopentyl_acetate", "phenethyl_acetate", "isobutyl_acetate",
                                             "ethanol..v.v.", "isoamyl_alcohol",
                                             "lactic_acid.mg.L.", "acetic_acid.g.L.")),
                     tibble(plotgroup = 'group 2',
                            field_name_R = c("bitterness", "alpha.humulene", "X1.alpha.terpineol", "citronellol",
                                             "geraniol", "linalool", "X2.3.butanedione", "X4VG", "acetaldehyde",
                                             "methylpyrazine")),
                     tibble(plotgroup = 'group 3',
                            field_name_R = c("ammonia.mg.L.", "iron", "DMS", "SO2.mg.L.", "glycerol", "GFS", 
                                             "protein.g.L.", "pH", "color.EBC.", "kcalperc"))) %>% 
  left_join(thresholds %>% 
              select(field_name_R,
                     chemical_name,
                     unit,
                     threshold_median,
                     origin_paper),
            by = c('field_name_R' = 'field_name_R')) %>% 
  mutate(clean_name = paste0(chemical_name,"\n(",unit,")")) %>% 
  nest(.by = plotgroup)
```

Here we add the corresponding chemical measurements from the beer dataset (stored in chem_calibrated) to each group of our plot.\
The data is in long format, so we have one column (property) that contains the 10 different chemicals within each group.

```{r}
rugplot_data <- rugplot_data %>% 
  mutate(beer_data = map(data,
                         .f = function(data){
                           beer_data = chem_calibrated %>% 
                             select(tasting_category_fine, data %>% pull(field_name_R)) %>% 
                             mutate(tasting_category_fine = gsub('Low/No alcohol', 
                                                                 'Non/low-alcohol', 
                                                                 tasting_category_fine)) %>% 
                             group_by(tasting_category_fine) %>%
                             pivot_longer(data %>% pull(field_name_R),
                                          names_to = "property",
                                          values_to = "measurement") %>%
                             left_join(data, by = c("property" = "field_name_R")) %>%
                             mutate(chemical_name = factor(chemical_name, levels = data %>% pull(chemical_name))) %>% 
                             ungroup()
                           return(beer_data)
                           })
         )
```

Creating a plotting function for the histograms and for the rugplots, then running them on each plotgroup, adding them to our dataframe.

```{r}
hist_plot = function(data){
  ggplot(data) +
    geom_rect(data = data %>%
                distinct(chemical_name, threshold_median),
              aes(xmin = 0, xmax = threshold_median, ymin = 0, ymax = 100),
              fill = "snow2", color = NA, alpha = 0.6) +
    geom_histogram(aes(measurement, fill = origin_paper), bins = 20, color = NA,
                   alpha = 1) +
    facet_grid(cols = vars(chemical_name), scales = "free") +
    scale_x_log10(labels = scales::label_log(base = 10, digits = 1),
                  breaks = scales::log_breaks(n = 4, base = 10)) +
    scale_y_continuous(name = "All styles", expand = c(0,0), limits = c(0,100)) +
    scale_color_manual(values = c("Malt" = color_malt, "Hops" = color_hop, 
                                  "Yeast" = color_yeast, "Wild flora" = color_wildflora, 
                                  "Others" = color_other, "Spices" = color_other)) +
    scale_fill_manual(values = c("Malt" = color_malt, "Hops" = color_hop, 
                                 "Yeast" = color_yeast, "Wild flora" = color_wildflora, 
                                 "Others" = color_other, "Spices" = color_other)) + 
    theme(strip.text.x = element_text(size = 6, margin = grid::unit(c(0,0,1,0), "mm")),
          plot.margin = grid::unit(c(1,1,0,6.3), "mm"),
          panel.background = element_blank(),
          panel.grid = element_blank(),
          panel.border = element_rect(colour = "snow2", fill = NA, linewidth = 0.5),
          panel.spacing = grid::unit(1, "points"),
          strip.background = element_blank(),
          axis.title.x = element_blank(),
          axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          axis.title.y = element_text(vjust = 0.5, hjust = 1, angle = 0, size = 7),
          axis.text.y = element_text(size = 5),
          legend.position = "none")
}

rugplot = function(data){
  ggplot(data) +
    geom_rect(data = data %>%
                distinct(chemical_name, threshold_median),
              aes(xmin = 0, xmax = threshold_median, ymin = 0, ymax = 1),
              fill = "snow2", color = NA, alpha = 0.6) +
    geom_vline(aes(xintercept = measurement, color = origin_paper),
               alpha = 0.7, size = 1) +
    facet_grid(rows = vars(tasting_category_fine), cols = vars(chemical_name), 
               scales = "free_x", switch = "both", 
               labeller = labeller(chemical_name = data %>% 
                                     distinct(chemical_name, unit) %>% 
                                     pull(unit, name = chemical_name))) +
    scale_x_log10(labels = scales::label_log(base = 10, digits = 1),
                  breaks = scales::log_breaks(n = 4, base = 10)) +
    scale_y_continuous(expand = c(0,0), limits = c(0,1), labels = NULL) +
    scale_color_manual(values = c("Malt" = color_malt, "Hops" = color_hop, 
                                  "Yeast" = color_yeast, "Wild flora" = color_wildflora, 
                                  "Others" = color_other, "Spices" = color_other)) +
    theme(strip.text.x = element_text(size = 6, vjust = 1, margin = margin(0,0,0,0, "mm")),
          strip.placement = "outside",
          strip.text.y.left = element_text(angle = 0, hjust = 1, vjust = 0.5, size = 6, 
                                           margin = grid::unit(c(0,1.4,0,0), "mm")),
          panel.background = element_blank(),
          panel.grid = element_blank(),
          panel.border = element_rect(colour = "snow2", fill = NA, linewidth = 0.5),
          panel.spacing = grid::unit(1, "points"),
          strip.background = element_blank(),
          plot.margin = grid::unit(c(0,1,0,1), "mm"),
          axis.title.x = element_blank(),
          axis.text.x = element_text(size = 6, margin = margin(0,0,0,0, "mm"), angle = 0),
          axis.ticks.y = element_blank(),
          legend.position = "none")
}

rugplot_data <- rugplot_data %>% 
  mutate(histograms = map(beer_data,
                          hist_plot),
         rugplots = map(beer_data,
                        rugplot))
```

```{r}
g_chem <- ggarrange(rugplot_data$histograms[[1]], rugplot_data$rugplots[[1]],
                    rugplot_data$histograms[[2]], rugplot_data$rugplots[[2]],
                    rugplot_data$histograms[[3]], rugplot_data$rugplots[[3]],
                    heights = c(1.2,4,1.2,4,1.2,4), 
                    nrow = 6, ncol = 1,
                    align = 'v')
```

```{r}
ggsave(g_chem, 
       filename = "chemical_data_rugplot.png",
       dpi = 600, width = 7.6, height = 9)
```

### Sensory rugplots

```{r}
attribute_origin = tibble(property = c("F_malt_all", "F_hops_all", "F_esters_all", "bitternes", "sweetness", 
                                       "acidity", "alcohol",  "body", "overall"),
                          clean_name = c("Malt taste", "Hop taste", "Ester taste", "Acidity", 
                                         "Bitterness", "Sweetness", "Alcohol", "Body fullness",
                                         "Overall\n(Trained panel)"),
                          color = c("color_malt", "color_hop", "color_yeast", "color_wildflora", "color_hop", 
                                    "color_malt", "color_yeast", "color_other", "color_other"))

sensory <- sensory_data %>%
  select(beer, beer_id, tasting_category_fine, F_malt_all, F_hops_all, F_esters_all, bitternes, sweetness, acidity, alcohol,  body, overall) %>% 
  pivot_longer(cols = c(-beer, -beer_id, -tasting_category_fine),
               names_to = 'property',
               values_to = 'measurement') %>% 
  left_join(attribute_origin,
            by = c('property' = 'property')) %>% 
  mutate(clean_name = factor(clean_name, levels = unique(clean_name))) %>% 
  mutate(tasting_category_fine = gsub('Low/No alcohol', 
                                      'Non/low-alcohol', 
                                      tasting_category_fine))


g_histogram_sens1 <- ggplot(sensory, aes(measurement)) +
  geom_histogram(aes(fill = color), bins = 15) +
  facet_grid(cols = vars(clean_name), scales = "free") +
  scale_y_continuous(name = "All styles", limits = c(0, 125)) +
  scale_fill_manual(values = c("color_malt" = color_malt, "color_hop" = color_hop, 
                               "color_yeast" = color_yeast, "color_wildflora" = color_wildflora, 
                               "color_other" = color_other)) + 
  theme_bw() +
  theme(strip.text.x = element_text(size = 6, margin = grid::unit(c(0,0,1,0), "mm")),
        plot.margin = grid::unit(c(1,1,0,6), "mm"),
        panel.background = element_blank(),
        panel.grid = element_blank(),
        panel.border = element_rect(colour = "grey", fill = NA, linewidth = 0.5),
        panel.spacing = grid::unit(1, "points"),
        strip.background = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.y = element_text(vjust = 0.5, hjust = 1, angle = 0, size = 7),
        axis.text.y = element_text(size = 5),
        legend.position = "none")


g_rugplot_sens1 <- ggplot(sensory, aes(x = measurement)) +
  geom_vline(aes(xintercept = measurement, color = color), alpha = 0.7, size = 1) +
  facet_grid(rows = vars(tasting_category_fine), cols = vars(clean_name), 
             scales = "free_x", switch = "both") +
  scale_y_continuous(expand = c(0,0), limits = c(0,1), labels = NULL) +
  scale_color_manual(values = c("color_malt" = color_malt, "color_hop" = color_hop, 
                                "color_yeast" = color_yeast, "color_wildflora" = color_wildflora, 
                                "color_other" = color_other)) + 
  theme_bw() +
  theme(strip.text.x = element_blank(),
        strip.placement = "outside",
        strip.text.y.left = element_text(angle = 0, hjust = 1, vjust = 0.5, size = 6, 
                                         margin = grid::unit(c(0,2.1,0,0), "mm")),
        panel.background = element_blank(),
        panel.grid = element_blank(),
        panel.border = element_rect(colour = "grey", fill = NA, size = 0.5),
        panel.spacing = grid::unit(1, "points"),
        strip.background = element_blank(),
        plot.margin = grid::unit(c(0,1,0,0), "mm"),
        axis.title.x = element_blank(),
        axis.text.x = element_text(size = 5, margin = margin(0,0,0,0, "mm"), angle = 0),
        axis.ticks.y = element_blank(),
        legend.position = "none")


g_sens <- ggarrange(g_histogram_sens1, g_rugplot_sens1,
                    heights = c(1.2,4), 
                    nrow = 2, ncol = 1)

ggsave(g_sens, 
       filename = "sensory_data_rugplot.png",
       dpi = 600, width = 7.6, height = 3)
```

Alternative, saving everything in one figure

```{r}
g_chem_sens <- ggarrange(rugplot_data$histograms[[1]], rugplot_data$rugplots[[1]],
                         rugplot_data$histograms[[2]], rugplot_data$rugplots[[2]],
                         rugplot_data$histograms[[3]], rugplot_data$rugplots[[3]],
                         g_histogram_sens1, g_rugplot_sens1,
                         heights = c(1.2,4,1.2,4,1.2,4,1.2,4), 
                         nrow = 8, ncol = 1)
ggsave(g_chem_sens, 
       filename = "chemical-sensory_subset_log_histogram_rugplot.png",
       dpi = 600, width = 7.6, height = 11)  
```

# Chemical and sensory correlation plots

Note: Because the RateBeer scores are property of RateBeer (ZX Ventures USA), only the data for the trained panel dataset is provided for this Figure.

Libraries

```{r}
library(tidyverse)
library(RColorBrewer)
library(gridExtra)
```

Colors

```{r}
color_yeast <- "#E86A3B"
color_malt <- "#6BA8C9"
color_hop <- "#ABD17D"
color_wildflora <- "#FACA13"
color_other <- "#5A5F62"
palette <- c(Malt = color_malt, 
             Hops = color_hop, 
             Yeast = color_yeast, 
             Spices = color_other, 
             Others = color_other,
             "Wild flora" = color_wildflora)
```

### Functions

Calculate the spearman correlation

```{r}
spearman_cor <- function(df){
  rho_cor <- cor(x = df %>% select(2), y = df %>% select(score), 
                 use = "pairwise.complete.obs", method = "spearman")
  return(c(df %>% distinct(cleanname), "rho" = format(round(rho_cor,2), nsmall = 2)))
}
```

Plotting function:

1.  Properly format the names of all variables
2.  Include threshold information
3.  Calculate spearman correlation with the function above

```{r}
chem_sense_facet <- function(data, chemical, sensory1, sensory2,
                             axis_title_size = 8,  axis_text_size = 7, padding = 0.02){
  chemical_name <- thresholds %>%
    filter(field_name_R == chemical) %>%
    mutate(compound_label = paste0(chemical_name," (",unit,")")) %>%
    pull(compound_label)
  sensory_name1 <- clean_names %>% 
    filter(R_name == sensory1) %>%
    pull(CleanName_PanelFeedback)
  sensory_name2 <- clean_names %>%
    filter(R_name == sensory2) %>%
    pull(CleanName_PanelFeedback)
  color <- palette[thresholds%>%
    filter(field_name_R == chemical) %>%
    pull(origin_paper)]
  
  threshold_median <- thresholds %>% filter(field_name_R == chemical) %>% pull(threshold_median)
  
  data_long <- data %>%
    select(beer, all_of(c(chemical, sensory1, sensory2))) %>% 
    pivot_longer(cols = c(sensory1, sensory2), names_to = 'aspect', values_to = 'score') %>% 
    left_join(data.frame(aspect = c(sensory1, sensory2), 
                         cleanname = c(sensory_name1, sensory_name2)), 
              by = c("aspect" = "aspect"))
  
  rho_spearman <- data_long %>%
    group_split(cleanname) %>%
    map(spearman_cor) %>%
    bind_rows() %>%
    mutate(!!sym(chemical) := 0.9*max(data_long %>% select(all_of(chemical)))) %>%
    mutate(score := max(data_long %>% select(score)))
  
  data_facet <- ggplot(data_long, aes(x = !!sym(chemical), y = score)) +
    annotate(xmin = -Inf, xmax = threshold_median, ymin = -Inf, ymax = Inf, geom = 'rect', 
             alpha = 0.6, fill = "snow2") +
    geom_point(size = 0.3, color = color, alpha = 0.8) +
    geom_text(data = rho_spearman, aes(label = paste0("ρ = ", rho)), size = 2.5,
              vjust = 1, hjust = 1) +
    scale_x_continuous(name = chemical_name, expand = c(padding, padding)) +
    facet_wrap(vars(cleanname), strip.position = "left", nrow = 2) +
    theme_bw() +
    theme(panel.grid = element_blank(),
          strip.background = element_blank(),
          strip.text.x = element_blank(),
          strip.text.y = element_text(size = axis_title_size, margin = margin(0,0,0,0)),
          strip.placement = "outside",
          axis.title = element_text(size = axis_title_size),
          axis.title.y = element_blank(),
          axis.text = element_text(size = axis_text_size),
          plot.margin=unit(c(2,2,2,2), "points"))
  
  return(data_facet)
}
```

```{r}
chem_sense_plot <- function(data, chemical, sensory, color, 
                            axis_title_size = 8,  axis_text_size = 7, padding = 0.02){
  chemical_name <- thresholds %>%
    filter(field_name_R == chemical) %>%
    mutate(compound_label = paste0(chemical_name," (",unit,")")) %>%
    pull(compound_label)
  sensory_name <- clean_names %>% 
    filter(R_name == sensory) %>%
    pull(CleanName_PanelFeedback)
  color <- palette[thresholds %>%
                     filter(field_name_R == chemical) %>%
                     mutate(origin_paper = as.character(origin_paper)) %>%
                     pull(origin_paper)]
  threshold_median <- thresholds %>% filter(field_name_R == chemical) %>% pull(threshold_median)
  
  data_long <- data %>%
    select(beer, !!sym(chemical), !!sym(sensory)) %>%
    gather(key = "aspect", value = "score", !!sym(sensory)) %>%
    left_join(data.frame(aspect = sensory, 
                         cleanname = sensory_name), 
              by = c("aspect" = "aspect"))
  
  rho_spearman <- data_long %>%
    group_split(cleanname) %>%
    map(spearman_cor) %>%
    bind_rows() %>%
    mutate(!!sym(chemical) := 0.9*max(data_long %>% select(!!sym(chemical)))) %>%
    mutate(score := max(data_long %>% select(score)))
  
  data_plot <- ggplot(data_long, aes(x = !!sym(chemical), y = score)) +
    annotate(xmin = -Inf, xmax = threshold_median, ymin = -Inf, ymax = Inf, geom = 'rect', 
             alpha = 0.6, fill = "snow2") +
    geom_point(size = 0.3, color = color, alpha = 0.8) +
    geom_text(data = rho_spearman, aes(label = paste0("ρ = ",rho)), size = 2.5,
              vjust = 1, hjust = 1) +
    scale_x_continuous(name = chemical_name, expand = c(padding, padding)) +
    scale_y_continuous(name = sensory_name) +
    theme_bw() +
    theme(panel.grid = element_blank(),
          strip.background = element_blank(),
          strip.text.x = element_blank(),
          strip.text.y = element_text(size = axis_title_size, margin = margin(0,0,0,0)),
          strip.placement = "outside",
          axis.title = element_text(size = axis_title_size),
          axis.text = element_text(size = axis_text_size),
          plot.margin=unit(c(2,2,2,2), "points"))
  
  return(data_plot)
}
```

### Gathering datasets

Chemical data calibrated

```{r}
chem_data = read.csv('Calibrated chemical dataset.csv')
sensory_data = read.csv("Supplemental File S4.csv")
thresholds <- read.csv("Sensory thresholds.csv")
clean_names <- read.csv("Sensory names.csv")
```

### Data pre-processing

```{r}
chem_calibrated = chem_data %>% 
  select(beer,
         ethyl_acetate, 
         ethyl_decanoate, 
         ethyl_hexanoate, 
         ethyl_octanoate,
         ethyl_propionate, 
         isobutyl_acetate, 
         isopentyl_acetate, 
         octyl_acetate, 
         phenethyl_acetate,
         ethyl_2.methyl_butyrate, 
         ethyl_isovalerate,
         ethanol..v.v., 
         isoamyl_alcohol, 
         isobutanol, 
         phenethyl_alcohol, 
         glycerol,
         bitterness, 
         alpha.humulene, 
         E.caryophyllene, 
         X1.alpha.terpineol, 
         citronellol, 
         geraniol, 
         linalool, 
         Z.nerolidol,
         acetic_acid.g.L., 
         lactic_acid.mg.L.,
         X2.3.butanedione, 
         X4VG, 
         acetaldehyde, 
         CS2, 
         DES, 
         DMS, 
         H2S, 
         MeSH, 
         SO2.mg.L.,
         CO2.PSI.,
         pH,
         ammonia.mg.L., 
         betaglucan.mg.L., 
         color.EBC.,
         iron, 
         protein.g.L.,
         GFS, 
         unfermentables, 
         kcalperc)

data_all  <- chem_calibrated %>% 
  left_join(sensory_data, by = c("beer" = "beer")) 
```

### Plotting correlations

```{r}
ethac_esters_facet <- chem_sense_facet(data_all,"ethyl_acetate", "A_esters_all","F_esters_all")
ethac_nailpolish_facet <- chem_sense_facet(data_all,"ethyl_acetate", "A_esters_ethac", "F_esters_ethac")
isoamyl_banana_facet <- chem_sense_facet(data_all,"isopentyl_acetate", "A_esters_isoaa", "F_esters_isoaa")
phenac_floral_facet <- chem_sense_facet(data_all,"phenethyl_acetate", "A_esters_flower", "F_esters_flower")
ethhex_fruity_facet <- chem_sense_facet(data_all,"ethyl_hexanoate", "A_esters_fruity", "F_esters_fruity")


ethanol_overall_plot <- chem_sense_plot(data_all,"ethanol..v.v.", "overall")
ethanol_alcohol_plot <- chem_sense_plot(data_all,"ethanol..v.v.", "alcohol")
isoalpha_bitter_plot <- chem_sense_plot(data_all,"bitterness", "bitternes")
gfs_sweet_plot <- chem_sense_plot(data_all,"GFS", "sweetness")
acetic_sourness_plot <- chem_sense_plot(data_all,"acetic_acid.g.L.", "acidity")


lay = rbind(c(1,2,3,4,5),
        c(1,2,3,4,5),
        c(6,7,8,9,10))

grid_graph <- grid.arrange(ethac_esters_facet, ethac_nailpolish_facet, 
                           isoamyl_banana_facet,
                           phenac_floral_facet, ethhex_fruity_facet,
                           ethanol_overall_plot, ethanol_alcohol_plot, 
                           acetic_sourness_plot, 
                           isoalpha_bitter_plot, gfs_sweet_plot, 
                           layout_matrix = lay) 

ggsave(grid_graph, 
       filename = "chemical_sensory_scatterplots.png",
       dpi = 600, width = 10, height = 4.5)
```

# Models trained with a style parameter

We reran the GBM model with a style parameter (actually one parameter per style, in a dummy-variable way)

```{r}
library(tidyverse)
```

#### Loading datasets

```{r}
Style_importance <- read.csv("GBM model with style information.csv")
clean_names <- read.csv("Sensory thresholds.csv")
```

#### Data wrangling

```{r}
Style_rank <- Style_importance %>% 
  left_join(clean_names %>% select(field_name_R, chemical_name), 
            by = join_by("Compound" == "field_name_R")) %>% 
  arrange(-MDI) %>% 
  mutate(MDI_rank = as.numeric(row.names(.))) %>% 
  arrange(-SHAP) %>% 
  mutate(SHAP_rank = as.numeric(row.names(.))) %>% 
  mutate(chemical_name = ifelse(is.na(chemical_name), 
                                gsub("_", " ", Compound),
                                chemical_name))


write.csv(Style_rank %>% filter(grepl("Style", chemical_name)),
          "Ranked importance of the different styles.csv",
          row.names = F)



MDI_style <- Style_rank %>% 
  rename(Importance = MDI) %>% 
  arrange(Importance) %>% 
  slice_tail(n = 15) %>% 
  mutate(chemical_name = ifelse(is.na(chemical_name), 
                                gsub("_", " ", Compound),
                                chemical_name),
         chemical_name = fct_inorder(chemical_name))
  

SHAP_style <- Style_rank %>% 
  rename(Importance = SHAP) %>% 
  arrange(Importance) %>% 
  slice_tail(n = 15) %>% 
  mutate(chemical_name = ifelse(is.na(chemical_name), 
                                gsub("_", " ", Compound),
                                chemical_name),
         chemical_name = fct_inorder(chemical_name))
```

#### Plotting

```{r}
{
  y_ax <- 'concentration'
  y_angle <- 90
  x_ax <- 'area'
  viridis_code = 'C'
  textsize = 12
  pointsize = 1.5
  legend_size = 0.85
  textcolor = '#192945'
}

{
  blue         = "#2c3e50" # blue
  red          = "#e31a1c" # red
  green        = "#18BC9C" # green
  yellow       = "#CCBE93" # yellow
  steel_blue   = "#a6cee3" # steel_blue
  navy_blue    = "#1f78b4" # navy_blue
  light_green  = "#b2df8a" # light_green
  pink         = "#fb9a99" # pink
  light_orange = "#fdbf6f" # light_orange
  orange       = "#ff7f00" # orange
  light_purple = "#cab2d6" # light_purple
  purple       = "#6a3d9a" # purple
}


plot_MDI_style <- ggplot(MDI_style, aes(y = chemical_name, x = Importance)) + 
  geom_bar(stat = "identity", fill = navy_blue) + 
  theme_tq() +
  labs(x = "MDI importance",
       y = "") +
  theme(axis.title.y = element_text(angle = y_angle, size = round(18*textsize/15), color = textcolor),
        axis.title.x = element_text(size = round(16*textsize/15), color = textcolor),
        axis.text.x = element_text(size = textsize, color = textcolor, angle = 0),
        axis.text.y = element_text(size = textsize, color = textcolor),
        axis.ticks = element_line(color = textcolor),
        plot.title = element_text(size = round(18*textsize/15)),
        legend.title = element_text(size = round(16*textsize/15), color = textcolor),
        legend.text = element_text(size = round(16*textsize/15), color = textcolor),
        legend.key.size = unit(legend_size, 'cm'),
        panel.grid = element_blank(),
        panel.background = element_rect(color = textcolor),
        strip.text = element_text(size = round(16*textsize/15), face = 'bold')) +
  ggtitle("") +
  lims(x = c(0, NA))


  

plot_SHAP_style <- ggplot(SHAP_style, aes(y = chemical_name, x = Importance)) + 
  geom_bar(stat = "identity", fill = orange) + 
  theme_tq() +
  labs(x = "SHAP importance",
       y = "") +
  theme(axis.title.y = element_text(angle = y_angle, size = round(18*textsize/15), color = textcolor),
        axis.title.x = element_text(size = round(16*textsize/15), color = textcolor),
        axis.text.x = element_text(size = textsize, color = textcolor, angle = 0),
        axis.text.y = element_text(size = textsize, color = textcolor),
        axis.ticks = element_line(color = textcolor),
        plot.title = element_text(size = round(18*textsize/15)),
        legend.title = element_text(size = round(16*textsize/15), color = textcolor),
        legend.text = element_text(size = round(16*textsize/15), color = textcolor),
        legend.key.size = unit(legend_size, 'cm'),
        panel.grid = element_blank(),
        panel.background = element_rect(color = textcolor),
        strip.text = element_text(size = round(16*textsize/15), face = 'bold')) +
  ggtitle("") +
  lims(x = c(0, NA))



Style_merge <- ggarrange(plot_MDI_style,
                         plot_SHAP_style)

ggsave("Style parameter importance.png",
       plot = Style_merge,
       units = 'mm',
       height = 200,
       width = 300)
```

# 100 Iterations of models

```{r}
library(tidyverse)
library(patchwork)
library(tidyquant)
```

#### Loading datasets

```{r}
performance <- read.csv("Performance of 100 models.csv")
importance <- read.csv("100 models feature importance.csv")
clean_names <- read.csv("Sensory thresholds.csv")
```

Preparing datasets

```{r}
performance <- performance %>% 
  rename("Gradient Boosting" = GBM,
         "Random Forest" = RF,
         "Extra Trees" = ET) %>% 
  pivot_longer(cols = -Iteration) %>% 
  mutate(name = fct_inorder(name))


importance <- importance %>% 
  pivot_longer(cols = X0:X99, names_to = "Iteration") %>% 
  mutate(Iteration = as.numeric(gsub("X", "", Iteration))) %>% 
  nest(data = c(Iteration, value)) %>% 
  left_join(clean_names %>% select(field_name_R, chemical_name), 
            by = join_by(Compound == field_name_R)) %>% 
  mutate(chemical_name = ifelse(is.na(chemical_name), "Performance", chemical_name),
         Average = as.numeric(map(data, 
                                  ~mean(.x$value))),
         Standdev = as.numeric(map(data,
                                   ~sd(.x$value))),
         lin_model = map(data,
                     ~lm(value ~ 1, .x)),
         conf_low = as.numeric(map(lin_model,
                                   ~confint(.x)[1])),
         conf_high = as.numeric(map(lin_model,
                                    ~confint(.x)[2]))) %>% 
  arrange(-Average) %>% 
  mutate(Model = fct_inorder(Model),
         std_low = Average - Standdev,
         std_high = Average + Standdev) %>% 
  group_by(Model, Score) %>% 
  slice(1:15)
```

#### Plotting performance, MDI and SHAP importance

```{r}
{
  y_ax <- 'concentration'
  y_angle <- 90
  x_ax <- 'area'
  viridis_code = 'C'
  textsize = 12
  pointsize = 1.5
  legend_size = 0.85
  textcolor = '#192945'
}

{
  blue         = "#2c3e50" # blue
  red          = "#e31a1c" # red
  green        = "#18BC9C" # green
  yellow       = "#CCBE93" # yellow
  steel_blue   = "#a6cee3" # steel_blue
  navy_blue    = "#1f78b4" # navy_blue
  light_green  = "#b2df8a" # light_green
  pink         = "#fb9a99" # pink
  light_orange = "#fdbf6f" # light_orange
  orange       = "#ff7f00" # orange
  light_purple = "#cab2d6" # light_purple
  purple       = "#6a3d9a" # purple
}


#### Performance

plot_performance <- ggplot(performance, aes(x = value)) + 
  facet_wrap(~ name) + 
  geom_boxplot(color = textcolor) + 
  coord_flip() +
  theme_classic() + 
  scale_x_continuous(n.breaks = 6,
                     limits = c(0, 1)) +
  theme_tq() +
  labs(x = "Performance (R2)",
       y = "",
       title = " ") +
  theme(axis.title.y = element_text(angle = y_angle, size = round(18*textsize/15), color = textcolor),
        axis.title.x = element_text(size = round(18*textsize/15), color = textcolor),
        axis.text.y = element_text(size = textsize, color = textcolor),
        axis.text.x = element_blank(),
        axis.ticks.y = element_line(color = textcolor),
        axis.ticks.x = element_blank(),
        plot.title = element_text(size = round(18*textsize/15), hjust = 0.5),
        legend.title = element_text(size = round(16*textsize/15), color = textcolor),
        legend.text = element_text(size = round(16*textsize/15), color = textcolor),
        legend.key.size = unit(legend_size, 'cm'),
        panel.grid = element_blank(),
        panel.background = element_rect(color = textcolor),
        strip.text = element_text(size = round(16*textsize/15), face = 'bold'),
        panel.spacing = unit(2, 'lines'))



#### Importance plots

plot_function = function(data, Model, Score, max_score){
  y_axis = if(Model == "GBM"){
    element_text(size = textsize, color = textcolor)
    } else {
      element_blank()
    }
  y_axis_ticks = if(Model == "GBM"){
    element_line(color = textcolor)
    } else {
      element_blank()
    }
  
  mycolor = if(Score == 'MDI'){
   navy_blue 
  } else {
    orange
  }
  
  ggplot(data, aes(x = chemical_name, y = Average)) + 
    geom_point(color = mycolor, size = 2.5) + 
    geom_errorbar(aes(xmin = chemical_name, xmax = chemical_name, ymin = std_low, ymax = std_high),
                  width = 0.01, color = mycolor) + 
    theme_tq() +
    labs(x = "",
         y = ifelse(Model == "GBM", paste(Score, "importance"), "")) +
    theme(axis.title.y = element_text(angle = y_angle, size = round(18*textsize/15), color = textcolor),
          axis.title.x = element_text(size = round(16*textsize/15), color = textcolor),
          axis.text.x = element_text(size = textsize, color = textcolor, angle = 90, vjust = 0.5, 
                                     hjust = 1),
          axis.text.y = y_axis,
          axis.ticks.x = element_line(color = textcolor),
          axis.ticks.y = y_axis_ticks,
          plot.title = element_text(size = round(18*textsize/15)),
          legend.title = element_text(size = round(16*textsize/15), color = textcolor),
          legend.text = element_text(size = round(16*textsize/15), color = textcolor),
          legend.key.size = unit(legend_size, 'cm'),
          panel.grid = element_blank(),
          panel.background = element_rect(color = textcolor),
          strip.text = element_text(size = round(16*textsize/15), face = 'bold')) +
    ggtitle("") + 
    ylim(0, max_score)
}


importance_plots <- importance %>% 
  group_by(Score) %>% 
  mutate(max_score = max(std_high)) %>% 
  nest(data = Compound:std_high) %>% 
  mutate(data = map(data, ~ .x %>% mutate(chemical_name = fct_inorder(chemical_name))), 
         plots = pmap(.l = list(data, Model, Score, max_score), 
                      .f = plot_function))
```

### Merging plots into one figure

```{r}
my_awesome_plot <- plot_performance + 
  importance_plots$plots + 
  plot_layout(design = '111
                        246
                        357')

ggsave("Results from 100 iterations.svg",
       plot = my_awesome_plot,
       units = "mm",
       width = 250,
       height = 350)
```
